---
title: "Chi Square"
author: "Deanna Elliott"
date: "11/25/2021"
output: html_document
---

```{r setup, include=TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(here)
library(janitor)
library(broom)
```

## Read in data

```{r}

lizards <- read_csv(here("data", "lizards.csv"))
```

Getting counts of lizards by species

```{r}

lizards %>% 
  count(spp) %>% 
  arrange(-n)
```

## Make a subset

 - UTST lizards
 - Sites: IBPE (enclosure grassland), RABB (rabbit mesquite), EAST (east tarbush)
 - Tail condition is recorded (W or B)
 
```{r}

utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B", "W"))
```
 
## Find counts and proportions of lizards with broken or whole tails at each site

```{r}

utst %>% 
  count(site, tail)
```

Easier way, better format for chi square test (janitor::tabyl, janitor::adorn_percentages)

```{r}

tail_counts <- utst %>% 
  tabyl(site, tail)  #count

tail_prop <- tail_counts %>% 
  adorn_percentages() %>% # proportion
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()
```

## Convert to contingency table

Use 'column_to_rowname()'

```{r}

lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-square

Are site and tail breakage independent of each other? 

```{r}

lizard_x2 <- chisq.test(lizard_ct)

lizard_tidy <- tidy(lizard_x2)
```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r round(lizard_tidy$statistic, 2)`, *p* = `r round(lizard_tidy$p.value, 4)`).


.
